An Implementation of the Active Contours without Edges Model and the Logic Framework for Active Contours on Multi-channel Images

### 2010-12-13 trunk/cs870/project/Report/Paper.tex
final version

\usepackage{cvpr}
\usepackage{cite}
\usepackage{url}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{array}
\usepackage{color}
\usepackage{balance}

@@

\title{An Implementation of the Active Contours without Edges model and the Logic Framework for Active contours on multi-channel images}

\author{Karim Ali and Sarah Nadi\\
\{karim, snadi\}@cs.uwaterloo.ca \\
David R. Cheriton School of Computer Science\\
University of Waterloo\\
}

@@

\begin{abstract}
In this project, we provide an implementation for the Chan-Vese model for active contours without edges, and the Sandberg-Chan logic framework for active
contours on multi-channel images. The Chan-Vese model is a special case of the more general Mumford-Shah functional for segmentation and level sets. It differs
from other active contour modesl in that it does not depend on the gradient of the contour ($\phi$), therefore it is more capable of detecting objects whose
boundaries may not be defined by a gradient (e.g. blurry images, noisy images). The Sandberg-Chan model builds on top of the Chan-Vese model by allowing the
contour to evolve simultaneously on multiple channels. It also offers a set of logic operations that can be applied to those channels (union, intersection,
complement, and their combinations). We evaluate our implementation based on several criteria, such as: independence of initial curve position, successful
detection of holes, performance with blurred and noisy images, and response to scaling parameters.
\end{abstract}

\section{Introduction}
Image segmentation or boundary detection is a very important problem in the area of Image Processing, and has received a lot of attention in the past. The
classical Active Contour model (or Snakes) proposed by Kass et al.~\cite{kass1988snakes} was the first model to use the idea of energy minimization to attract
a contour to the edges of the objects in an image. The Snakes model was very successful and variations of it very highly adopted later on
(E.g.~\cite{caselles1997geodesic}. Most of these models used the level set formulation for propagating fronts to evolve the curve. However, these models highly
depended on curvature motion (motion defined by the gradient of the curve) which led to poor performance in smoothed edges. To overcome these limitations, Chan
and Vese~\cite{chan2001active} propose an active contour model that does not depend on the edges (i.e the gradient) for propagating the curve to detect the
boundary of the object. Instead, they use a region-based approach based on the Mumford-Shah model~\cite{mumford1989optimal} to divide the image
into two regions: one inside the propagating curve, and one outside. The curve is at the boundary of the object if there is no difference in intensities inside
the curve as well as outside the curve.

The Active Contours without Edges model proposed by Chan and Vese (referred to as Chan-Vese model throughout the rest of this paper) was very successful in
detecting objects even in noisy or blurry images. It could also detect holes in objects which was usually a limitation in previous models. As an extension to
this work, Sandberg and Chan~\cite{sandberg2005logic} propose a logic framework (referred to as the Sandberg-Chan model throughout the rest of this
paper) that  performs logical operations on multiple images according to the curve propagation proposed in the Chan-Vese model. To achieve that, previous models
usually had two steps. They would either first segment the object in each channel separately then combine the segmented objects according to the logic operation
through bitwise operations (CITE) or they would apply logic operations to the different images then segment the resulting image (CITE). The first approach is
very costly, and the second approach requires a lot of prior knowledge about the intensities of each image. To overcome these drawbacks, the Sandberg-Chan
model is based on the idea of fitting a single contour to the object on all channels according to the logic operator, and based on regions.

In this paper, we report on our findings after implementing each of these models. We implemented both models in Matlab, and experimented with several images.
In this paper, we explain the details of our implementation as well as our results. The rest of this paper is organized as follows: Section~\ref{sec:bg} first
provides brief background about each of the two models implemented in this paper. Section~\ref{sec:implementation} then explains our implementation. In this
section, we explain how we implemented the models and any variations from the original papers. Section~\ref{sec:results} explains our evaluation criteria, and
presents the results we obtained. Section~\ref{sec:difficulties} discusses some of the difficulties we faced, and points out some BLA. Section~\ref{sec:concl}
concludes this paper by summarizing our findings.

@@

The Chan-Vese model is a region based model for detecting objects in an image. It is based on a restriction Mumford-Shah model which divides an image into
regions and represents each region by a piecewise constant (the minimal partition problem). Figure~\ref{fig:region} shows what is meant by a region based
model. The figure shows an image $u_{0}$ which has a gray triangular object in it. The red curve, C, is the initial contour used to detect this object. The
main idea behind the model is that the curve divides the image into two regions: that inside the curve and that outside. Each region is represented by a
constant, c, which is the average intensity of the image values in each region. In order for the curve to fit the object, there must be no variation of the
intensities inside the curve as well as outside. In other words, this turns into a minimization problem of the difference of intensities inside (fitting term
one, $F_{1}$) plus those
outside (fitting term two, $F_{2}$). For example, in the figure, the point $(x_{1},y_{1})$ will have to be outside the curve in order to minimize the difference
between the points inside
the curve. Similarly, the point $(x_{4}, y_{4})$ will have to be inside the curve. Figure~\ref{fig:fitting} shows all the possibilities of the curve's fitting
according to its position with respect to the object.

More formally, the Euler-Lagrange equation (as derived in the original paper) representing the time motion of the curve C is shown in Equation~\ref{eqn:cvpde}.

@@

(REST OF EQUATION HERE)

The parameter $\mu$ is a scaling parameter for the length of the curve represented in terms of curvature. The smalled $\mu$ is, the more the length of the curve
can increase without penalizing the minimization. This allows the model to detect smalled objects and holes. The larger $\mu$ is, the less freedom there is for
the curve to increase in length, and thus, it will only be able to detect larger objects. The parameter $\nu$ is also a scaling term for the area of the curve.
However, the authors do not use the area term in the Euler-Lagrange derivation, and always set $\nu$ to 0. It seems that $\mu$ is sufficient to scale the curve
according to the objects that need to be detected. Finally, $\lambda_{1}$ and $\lambda_{2}$ are weighting parameters for the forces inside the curve and
outside the curve respectively. Since we want to give both forces equal weight, the authors set $\lambda_{1} = \lambda_{2} = 1$ in all their experiments.

@@

The main goal of the Sandberg-Chan model~\cite{sandberg2005logic} is to perform logic operations on a combination of different images accurately and
efficiently. For example, finding the union of two images where different parts of the object are occluded in each image so that a complete object can be
obtained. In order to do that, they use the Chan-Vese model to detect objects, and simultaneously include the logic operations in the minimization problem.
Since they are using the Chan-Vese model, the problem must be viewed in terms of regions as well. Accordingly, they define the two main logic operations
(union and intersection) in terms of regions where the union of two images is the union of the insides of the curve with respect to the images plus the
intersection of the outsides of the curve with respect to the images. For example, Figure~\ref{fig:logic-op} (taken from the original paper) shows that the
union of $A_{1}$ and $A_{2}$ (in terms of its shape) can be obtained by taking the union of the insides of the object or the intersection of the outsides.
Similarly, the intersection of two images would be the intersection of the insides plus the union of the outsides. Thus, in order to obtain accurate results
irrespective of the object's intensity outside versus inside, we should consider \textit{both} the logical operation required for the region inside the object
as well as that outside.

More formally, the authors define two logic variables $z_{i}^{in}$ and $z_{i}^{out}$ to denote whether a point $(x,y)$ should be in the moving curve C or not
with respect to image i. Since we are trying to minimize the fitting of the curve, they use 0 to denote true and 1 to denote false (the reverse of the usual
convention. Following the Chan-Vese model, they represent each of the regions inside and outside the curve by a constant c. In this paper, they represent
$c_{in}$ as $c_{+}$ and $c_{out}$ as $c_{-}$. Equations~\ref{eqn:zin} and~\ref{eqn:zout} show how to calculate $z_{in}$ and $z_{out}$ respectively in terms of
the Chan-Vese model. We note here that in the original paper there was a typo in these equations where they divide by the maximum intensity of each image
instead of the maximum intensity squared. However, in order to have $z_{in}$ and $z_{out}$ have values from 0 to 1 that represent logic values, we need to
divide by the maximum intensity squared. The equations shown below have been corrected for that.

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{logicop.png}
\caption{Region Based View Union}
\label{fig:logic-op}
\end{figure*}

@@

\begin{equation}
\label{eqn:finters}
f_{\cap} = 1 = ((1 - z_{1}) .(1-  z_{2}))^{1/2}
\end{equation}

Based on these functions, the Euler-Lagrange equation for any logic operation is shown in Equation~\ref{eqn:svpde}. According to the desired logic operation,
$f_{in}$ and $f_{out}$ will be specified accordingly. For example, if we are doing the union of the images, then we will need to perform a union operation on
the insides. Thus, $f_{in}$ will be replaced by Equation~\ref{eqn:funion}. Similarly, we will need to do the intersection of the outsides so $f_{out}$ will be
replaced by Equation~\ref{eqn:finters}.

@@

During implementation, we tried to stick to the authors' formulas and guidelines as much as possible. In this section we explain the choices we made regardless
implementation. For the chan-vese model, we used the PDE given in Equation 9 in the original paper~\cite{chan2001active} which is shown in
Equation~\ref{eqn:cvpde}.  We, first, tried to implement the delta dirac function, $\delta_{\varepsilon}\phi$, as defined in the paper. However, we did not
get non-zero values everywhere as indicated by the authors. Accordingly, we chose to use $|\bigtriangledown\phi|$ instead of the delta dirac function as this
was indicated as a valid alternative by the authors. To calculate $c_{1}$ and $c_{2}$, we simply calculated the mean of the values inside $\phi$ (specifically
where $\phi > 0$) and the mean of the values outside $\phi$ (specifically where $\phi < 0$) respectively. To actually solve the PDE, we chose to use an explicit
time stepping scheme for the finite difference discretization. That is, in each time step $\phi_{t} = \phi + \triangle t*force$. This was simpler to implement,
and although $\triangle t$ must be chosen carefully to satisfy the stability condition, we did not suffer from performance problems due to this restriction.

For the sandberg-chan model, we used the PDE given in original paper as well, and shown in Equation~\ref{eqn:svpde}. Again, we used $|\bigtriangledown\phi|$
instead of the dirac function. We calculated $z_{in}$ and $z_{out}$ according to Equations~\ref{eqn:zin, eqn:zout} shown above which have been corrected for
the typo in the original paper. For all other equations and calculations necessary in the model, we closely followed the original paper in our implementation.
We also used an explicit time stepping scheme in this model.

@@

In order to make sure we correctly implemented the models, we had several evaluation criteria. Table~\ref{tab:criteria} summarizes these criteria, and explains
the goals of each. Some of these criteria apply for both models, while other apply to one or the other. For the rest of this section, we proceed by presenting
our experimental results for both models. For each model, we proceed in the order of these criteria starting with the simplest
cases, and incrementally challenging the model. Unless otherwise stated, all the segmentation was performed on a gray scale version of the original image. All
the images in the Chan-Vese model were run on a BLA COMPUTER and all the images in the Sandberg-Chan model were run on a BLA computer. As shown in the results,
we tried different image sizes to ensure that our solution converges.

@@

@@

Figure~\ref{fig:cv_eg1} shows that the implemented model can successfully detect object boundaries in a simple image. In this case, the evolving curve was
successfully able to detect the boundaries of both objects. Figure~\ref{fig:cv_eg2} shows a slightly more complicated example where the evolving contour
successfully detects the contour of the brain. Note that there is a tiny area to the bottom left of the brain that was also successfully detected. To ensure
that open boundaries (e.g. lines) can also be detected, we used an image with several shapes shown in Figure~\ref{fig:cv_eg8};

\begin{figure*}[t]
\centering
\includegraphics[width=11cm]{cv_eg1.png}
\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
0.01$, no reinitialization, cpu = 2.9s, 7 iterations.}
\label{fig:cv_eg1}
\end{figure*}

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg2.png}
\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.6)^2 + (y - 65.5)^2)} + 32.8$, $\mu =
0.01$, no reinitialization, cpu = 1.95 s, took 7 iterations.}
\label{fig:cv_eg2}
\end{figure*}

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg8.png}
\caption{Successful detection of object with open boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
0.01$, no reinitialization, cpu = 5.19 s, took 11 iterations.}
\label{fig:cv_eg8}
\end{figure*}

@@

To ensure that the position of the curve does not affect the final segmentation, we tested two other positions for the initial curve for the same brain image
used in Figure~\ref{fig:cv_eg2}. In Figure~\ref{fig:cv_eg2}, the initial contour was completely overlapping the object. Accordingly, we tried two other
positions. The first one is shown in Figure~\ref{fig:cv_eg3} where the initial contour only partially overlaps with the object. The second one is shown in
Figure~\ref{fig:cv_eg4} where the initial contour does not overlap the object in any part. Visually, the obtained segmentation was the same with a slightly
higher number of iterations and CPU time. In all three cases, the brain was correctly segmented. We compared the area segmented as the brain in each case to
make sure the same segmentation was obtained, and in each case, we got the same area of 0.5268. Note, however, that when part of the curve lies outside the
object, the definition of inside and outside changes since the curve becomes an open curve at one point. This should not make a difference as long as the
contour lies correctly at the boundaries of the object.

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg3.png}
\caption{Initial curve position partially overlapping object. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 30)^2 + (y - 30)^2)} + 20$, $\mu =
0.01$, no reinitialization, cpu = 2.26 s, took 9 iterations.}
\label{fig:cv_eg3}
\end{figure*}

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg4.png}
\caption{Initial curve position not overlapping any area of the object. Top: the evolving curve (in red) over time where the first image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2)} + 10$,
$\mu =0.01$, no reinitialization, cpu = 2.01 s, took 9 iterations.}
\label{fig:cv_eg4}
\end{figure*}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg5.png}
\caption{Successful detection of holes.  Top: the evolving curve (in red) over time where the first image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.5)^2 + (y - 65.5)^2)} + 32.8$,
$\mu =0.01$, no reinitialization, cpu = 8.12 s, took 12 iterations.}
\label{fig:cv_eg5}
\end{figure*}

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg6.png}
\caption{Successful detection of multiple holes.  Top: the evolving curve (in red) over time where the first image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
$\mu =0.01$, no reinitialization, cpu = 15.07 s, took 23 iterations.}
\label{fig:cv_eg6}
\end{figure*}

\begin{figure*}[t]
\centering
\includegraphics[width=10cm]{cv_eg7.png}
\caption{Inability to detect low intensities that have not much variation from their background.  Top: the evolving curve (in red) over time where the first
image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
$\mu =0.01$, no reinitialization, cpu = 1.35 s, took 6 iterations.}
\label{fig:cv_eg7}
\end{figure*}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg9.png}
\caption{Ability to detect contours in blurry images.  Top: the evolving curve (in red) over time where the first
image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 255 x 255, $\phi_{0}(x,y) = - \sqrt{(x - 127.5)^2 + (y - 127.5)^2)} +
63.75$, $\mu =0.01$, no reinitialization, cpu = 5.41 s, took 15 iterations.}
\label{fig:cv_eg9}
\end{figure*}

@@

For the Sandberg-Chan model, the successful detection of boundaries criteria is implicitly included in the logic operations functionality. We therefore, show
the other criteria. Unless otherwise specified, $\mu = 0.1$ and $\lambda = 255*255$ in the experiments below.

>>>>

\usepackage{cvpr}
\usepackage{times}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{url}
\usepackage{appendix}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi

@@

\title{An Implementation of the Active Contours without Edges Model and the Logic Framework for Active Contours on Multi-channel Images}

\author{Karim Ali and Sarah Nadi\\
\{karim, snadi\}@cs.uwaterloo.ca \\
CS870 - Fall 2010\\
David R. Cheriton School of Computer Science\\
University of Waterloo\\
}

@@

\begin{abstract}
In this project, we provide an implementation for the Chan-Vese model for active contours without edges, and the Sandberg-Chan logic framework for active
contours on multi-channel images. The Chan-Vese model is a special case of the more general Mumford-Shah functional for segmentation and level sets. It
differs
from other active contour models in that it is not edge dependent, therefore it is more capable of detecting objects whose
boundaries may not be defined by a gradient (e.g. blurry images, noisy images). The Sandberg-Chan model builds on top of the Chan-Vese model by allowing the
contour to evolve simultaneously on multiple channels. It also offers a set of logic operations that can be applied to those channels (union, intersection,
complement, and their combinations). We evaluate our implementation based on several criteria, such as: independence of initial curve position, successful
detection of holes, performance with blurred and noisy images, and response to scaling parameters.
\end{abstract}

\section{Introduction}
Image segmentation or boundary detection is a very important problem in the area of Image Processing, and has received a lot of attention in the past. The
classical Active Contour model (or Snakes) proposed by Kass et al.~\cite{kass1988snakes} was the first model to use the idea of energy minimization to attract
a contour to the edges of the objects in an image. The Snakes model was very successful and variations of it were highly adopted later on
(E.g.~\cite{caselles1997geodesic}. Most of these models use the level set formulation for propagating fronts to evolve the curve. However, these models highly
depend on motion defined by the gradient of the image which leads to poor performance in smoothed edges. To overcome these limitations, Chan
and Vese~\cite{chan2001active} propose an active contour model that does not depend on the edges (i.e the gradient) for propagating the curve to detect the
boundary of the object. Instead, they use a region-based approach based on the Mumford-Shah model \cite{mumford1989optimal} to divide the image
into two regions: one inside the propagating curve, and one outside. The curve is at the boundary of the object if there is no difference in intensities inside
the curve as well as outside the curve.

The Active Contours without Edges model proposed by Chan and Vese (referred to as Chan-Vese model throughout the rest of this paper) is very successful in
detecting objects even in noisy or blurry images. It can also detect holes in objects which is usually a limitation in previous models. As an extension to
this work, Sandberg and Chan~\cite{sandberg2005logic} propose a logic framework (referred to as the Sandberg-Chan model throughout the rest of this
paper) that  performs logic operations on multiple images according to the curve propagation proposed in the Chan-Vese model. To achieve that, previous models
usually had two steps. They would either first segment the object in each channel separately then combine the segmented objects according to the logic operation
through bitwise operations~\cite{sapiro1997color, zhu2002region} or they would apply logic operations to the different images then segment the resulting
image. The first approach is
very costly, and the second approach requires a lot of prior knowledge about the intensities of each image. To overcome these drawbacks, the Sandberg-Chan
model is based on the idea of fitting a single contour to the object on all channels according to the logic operator, and based on regions.

In this paper, we report on our findings after implementing each of these models. We implement both models in Matlab, and experiment with several images.
We explain the details of our implementation as well as our results. The rest of this paper is organized as follows:
Section~\ref{sec:bg} provides brief background about each of the two models implemented in this paper. Section~\ref{sec:implementation} then explains our
implementation. In this section, we explain how we implemented the models and any variations from the original papers. Section~\ref{sec:results} explains our
evaluation criteria, and
presents the results we obtained. Section~\ref{sec:difficulties} discusses some of the difficulties we faced during implementation. Section~\ref{sec:concl}
concludes this paper by summarizing our findings.

@@

The Chan-Vese model~\cite{chan2001active} is a region based model for detecting objects in an image. It is based on a restriction of the Mumford-Shah model
which divides an image into
regions and represents each region by a piecewise constant (also called the minimal partition problem). Figure~\ref{fig:region} shows this region based model.
The figure shows an image $u_{0}$ which has a gray triangular object in it. The red curve, $C$, is the initial contour used to detect this object. The
main idea behind the model is that the curve divides the image into two regions: that inside the curve and that outside. Each region is represented by a
constant, $c$, which is the average intensity of the image values in each region. In order for the curve to fit the object, there must be minimal variation of
the
intensities inside the curve as well as outside. In other words, this turns into a minimization problem of the difference of intensities inside (fitting term
one, $F_{1}$) plus those
outside (fitting term two, $F_{2}$). For example, in Figure~\ref{fig:region}, the point $(x_{1},y_{1})$ will have to be outside the curve in order to minimize
the difference
between the points inside
the curve. Similarly, the point $(x_{4}, y_{4})$ will have to be inside the curve. Figure~\ref{fig:fitting} shows all the possibilities of the curve's fitting
according to its position with respect to the object.

More formally, the Euler-Lagrange equation (as derived in the original paper) representing the time motion of the curve $C$ is shown in
Equation~\ref{eqn:cvpde}.

@@

The parameter $\mu$ is a scaling parameter for the length of the curve represented in terms of curvature. The smaller $\mu$ is, the more the length of the curve
can increase without penalizing the minimization. This allows the model to detect smaller objects and holes. The larger $\mu$ is, the less freedom there is for
the curve to increase in length, and thus, it will only be able to detect larger objects. The parameter $\nu$ is also a scaling term for the area of the curve.
However, the authors do not use the area term in the Euler-Lagrange derivation, and always set $\nu$ to 0. It seems that $\mu$ is sufficient to scale the curve
according to the objects that need to be detected. Finally, $\lambda_{1}$ and $\lambda_{2}$ are weighting parameters for the forces inside the curve and
outside the curve respectively. Since we want to give both forces equal weight, the authors set $\lambda_{1} = \lambda_{2} = 1$ in all their experiments.

\begin{figure*}[t]
\centering
\includegraphics[width=0.7\textwidth]{logicop.png}
\caption{Region Based View Union}
\label{fig:logic-op}
\end{figure*}

@@

The main goal of the Sandberg-Chan model~\cite{sandberg2005logic} is to perform logic operations on a combination of different images accurately and
efficiently. For example, finding the union of two images where different parts of the object are occluded in each image so that a complete object can be
obtained. This example will be shown later in Figure~\ref{fig:family}. In order to do that, they use the Chan-Vese model to detect objects, and simultaneously
include the logic operations in the minimization problem.
Since they are using the Chan-Vese model, the problem must be viewed in terms of regions as well. Accordingly, they define the two main logic operations
(union and intersection) in terms of regions where the union of two images is the union of the inside of the curve with respect to the images plus the
intersection of the outside of the curve with respect to the images. For example, Figure~\ref{fig:logic-op}~\cite{sandberg2005logic} shows that the
union of $A_{1}$ and $A_{2}$ (in terms of its shape) can be obtained by taking the union of the insides of the object or the intersection of the outsides.
Similarly, the intersection of two images would be the intersection of the insides plus the union of the outsides. Thus, in order to obtain accurate results
irrespective of the object's intensity outside versus inside, we should consider \textit{both} the logic operation required for the region inside the object
as well as that outside.

More formally, the authors define two logic variables $z_{i}^{in}$ and $z_{i}^{out}$ to denote whether a point $(x,y)$ should be in the moving curve $C$ or not
with respect to image $i$. Since we are trying to minimize the fitting of the curve, they use $0$ to denote $true$ and $1$ to denote $false$ (the reverse of the
usual
convention). Following the Chan-Vese model, they represent each of the regions inside and outside the curve by a constant $c$. In this paper, they represent
$c_{in}$ as $c_{+}$ and $c_{out}$ as $c_{-}$. Equations~\ref{eqn:zin} and~\ref{eqn:zout} show how to calculate $z_{in}$ and $z_{out}$ respectively in terms of
the Chan-Vese model. We note here that in the original paper there was a typo in these equations where they divide by the maximum intensity of each image
instead of the maximum intensity squared. However, in order to have $z_{in}$ and $z_{out}$ have values from 0 to 1 that represent logic values, we need to
divide by the maximum intensity squared. The equations shown below have been corrected for that.

@@

\begin{equation}
\label{eqn:finters}
f_{\cap} = 1 - ((1 - z_{1}) .(1-  z_{2}))^{1/2}
\end{equation}

Based on these functions, the Euler-Lagrange equation for any logic operation is shown in Equation~\ref{eqn:svpde}. According to the desired logic operation,
$f_{in}$ and $f_{out}$ will be specified. For example, if we are doing the union of the images, then we will need to perform a union operation on
the insides. Thus, $f_{in}$ will be replaced by Equation~\ref{eqn:funion}. Similarly, we will need to do the intersection of the outsides so $f_{out}$ will be
replaced by Equation~\ref{eqn:finters}.

@@

During implementation, we tried to stick to the authors' formulas and guidelines as much as possible. In this section we explain what methods we use to
numerically solve the PDEs. For the chan-vese model, we use the PDE given in Equation 9 in the original paper~\cite{chan2001active} which is shown in
Equation~\ref{eqn:cvpde}.  We, first, tried to implement the dirac delta function, $\delta_{\varepsilon}\phi$, as defined in the paper. However, it resulted in
worse segmentations. Accordingly, we chose to use $|\bigtriangledown\phi|$ instead of the dirac delta function as this
was indicated as a valid alternative by the authors. To calculate $c_{1}$ and $c_{2}$, we simply calculated the mean of the values inside $\phi$ (specifically
where $\phi > 0$) and the mean of the values outside $\phi$ (specifically where $\phi < 0$) respectively. To actually solve the PDE, we chose to use an explicit
time stepping scheme for the finite difference discretization. That is, in each time step $\phi^{n+1} = \phi^n + \triangle t*phi_t$. This was simpler to
implement, and although $\triangle t$ must be chosen carefully to satisfy the stability condition ($\triangle t <= (1/imageSize)^2$), we did not suffer from
performance problems due to this restriction.

For the Sandberg-Chan model, we used the PDE given in original paper as well, and shown in Equation~\ref{eqn:svpde}. Again, we used $|\bigtriangledown\phi|$
instead of the dirac delta function. We calculated $z_{in}$ and $z_{out}$ according to Equations~\ref{eqn:zin} and~\ref{eqn:zout} shown above which have been
corrected for
the typo in the original paper. For all other equations and calculations necessary in the model, we closely followed the original paper in our implementation.
We also used an explicit time stepping scheme in this model.

We use the same stopping condition for both model. That is, to check if we have reached the stationary solution, we check if more points have been added inside
the contour or not. More formally, we check if $(\phi^{n+1} > 0) == (\phi_{n} > 0)$. If the contour did not move, then new points will be added to its inside,
and this condition will be satisfied.

@@

In order to make sure we correctly implemented the models, we had several evaluation criteria. Table~\ref{tab:criteria} summarizes these criteria, and explains
the goals of each. Some of these criteria apply for both models, while other apply to one or the other. For the rest of this section, we proceed by presenting
our experimental results for both models. For each model, we proceed in the order of these criteria starting with the simplest
cases, and incrementally challenging the model. Unless otherwise stated, all the segmentation was performed on a gray scale version of the original image. All
the images in the Chan-Vese model were run on an Intel Core 2 Duo processor with 2.00 GHz and a 2GB RAM and all the images in the Sandberg-Chan
model were run on a Intel Centrino Duo 2.0 processor and a 1GB RAM computer. As shown in the results,
we tried different image sizes to ensure that our solution converges.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{cv_eg1.png}
\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$, $\mu =
0.01$, no reinitialization, cpu = 2.9s, iterations = 7.}
\label{fig:cv_eg1}
\end{figure}

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg2.png}
\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.6)^2 + (y - 65.5)^2} + 32.8$, $\mu =
0.01$, no reinitialization, cpu = 1.95 s, iterations = 7.}
\label{fig:cv_eg2}
\end{figure*}

@@

Figure~\ref{fig:cv_eg1} shows that the implemented model can successfully detect object boundaries in a simple image. In this case, the evolving curve was
successfully able to detect the boundaries of both objects. Figure~\ref{fig:cv_eg2} shows a slightly more complicated example where the evolving contour
successfully detects the contour of the brain. Note that there is a tiny area near the center to the left of the brain that was also successfully detected. To
ensure
that open boundaries (e.g. lines) can also be detected, we used an image with several shapes shown in Figure~\ref{fig:cv_eg8}.

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg8.png}
\caption{Successful detection of object with open boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until all objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$, $\mu =
0.01$, no reinitialization, cpu = 5.19 s, iterations = 11.}
\label{fig:cv_eg8}
\end{figure*}

\begin{figure*}[t!]
\centering
\includegraphics[width=12cm]{cv_eg3.png}
\caption{Initial curve position partially overlapping object. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 30)^2 + (y - 30)^2} + 20$, $\mu =
0.01$, no reinitialization, cpu = 2.26 s, iterations = 9.}
\label{fig:cv_eg3}
\end{figure*}

\begin{figure*}[t!]
\centering
\includegraphics[width=12cm]{cv_eg4.png}
\caption{Initial curve position not overlapping any area of the object. Top: the evolving curve (in red) over time where the first image shows the initial
contour. Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2} + 10$,
$\mu =0.01$, no reinitialization, cpu = 2.01 s, iterations = 9.}
\label{fig:cv_eg4}
\end{figure*}

@@

To ensure that the initial position of the curve does not affect the final segmentation, we tested two other positions for the initial curve for the same
brain image
used in Figure~\ref{fig:cv_eg2}. In Figure~\ref{fig:cv_eg2}, the initial contour was completely overlapping the object. Accordingly, we tried two other
positions. The first one is shown in Figure~\ref{fig:cv_eg3} where the initial contour only partially overlaps with the object. The second one is shown in
Figure~\ref{fig:cv_eg4} where the initial contour does not overlap the object in any part. Visually, the obtained segmentation was the same with a slightly
higher number of iterations and CPU time. In all three cases, the brain was correctly segmented. We compared the area segmented as the brain in each case to
make sure the same segmentation was obtained, and in each case, we got the same area of 0.5268. Note, however, that when part of the curve lies outside the
object, the definition of inside and outside changes since the curve becomes an open curve at one point. This should not make a difference as long as the
contour lies correctly at the boundaries of the object.

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg5.png}
\caption{Successful detection of holes.  Top: the evolving curve (in red) over time where the first image shows the initial
contour. Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.5)^2 + (y - 65.5)^2} + 32.8$,
$\mu =0.01$, no reinitialization, cpu = 8.12 s, iterations = 12.}
\label{fig:cv_eg5}
\end{figure*}

\begin{figure*}[t!h!]
\centering
\includegraphics[width=12cm]{cv_eg6.png}
\caption{Successful detection of multiple holes.  Top: the evolving curve (in red) over time where the first image shows the initial
contour. Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$,
$\mu =0.01$, no reinitialization, cpu = 15.07 s, iterations =  23.}
\label{fig:cv_eg6}
\end{figure*}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{cv_eg7.png}
\caption{Inability to detect low intensities that do not have much variation from their background.  Top: the evolving curve (in red) over time where the
first
image shows the initial
contour. Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$,
$\mu =0.01$, no reinitialization, cpu = 1.35 s, iterations = 6.}
\label{fig:cv_eg7}
\end{figure}

@@

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=12cm]{cv_eg9.png}
\caption{Ability to detect contours in blurry images.  Top: the evolving curve (in red) over time where the first
image shows the initial
contour. Bottom: evolving segmentation over time until the object is detected. Size = 255 x 255, $\phi_{0}(x,y) = - \sqrt{(x - 127.5)^2 + (y - 127.5)^2} +
63.75$, $\mu =0.01$, no reinitialization, cpu = 5.41 s, iterations = 15.}
\label{fig:cv_eg9}
\end{figure*}

@@

For the Sandberg-Chan model, the successful detection of boundaries criterion is implicitly included in the logic operations functionality. We therefore, show
the other criteria. Unless otherwise specified, $\mu = 0.1$ and $\lambda = 255*255$ in the experiments below.

\begin{figure*}[t!]
\centering
\includegraphics[width=0.5\textwidth]{cv_eg12.png}
\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the final segmentation in each case.}
\label{fig:cv_eg12}
\end{figure*}
